---
title: "Code_Handout"
author: "Jackie Vogel, Sarah Millard, Ignacio Luque, Leo Walker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# installing caret for train/test split
library(caret)
library(nnet)
library(tidyverse)
library(lmtest) # For lrtest()
library(caret)
```

## Multinomial Logistic Regression

This R Markdown demonstrates how to use Multinomial Logistic Regression on a small data set that is predicting which drug to prescribe to a patient.

Data set was found at https://www.kaggle.com/datasets/jeevanrh/drug200csv


```{r}
drug <-read.csv("../data/drug200.csv", header=TRUE, sep= ",")
str(drug)
```

## Change the chr types to factors


```{r}
drug$Sex <- as.factor(drug$Sex)
drug$BP <- as.factor(drug$BP)
drug$Cholesterol <- as.factor(drug$Cholesterol)
drug$Drug <- as.factor(drug$Drug)

str(drug)
```
##Next, have a look at each variable in the data set.

```{r}

table(drug$Drug, exclude = NULL)

table(drug$Na_to_K, exclude = NULL)

table(drug$Cholesterol, exclude = NULL)

table(drug$BP, exclude = NULL)

table(drug$Sex, exclude = NULL)

table(drug$Age, exclude = NULL)

```

##Model 1

```{r}

Model.1 <- glm(Drug ~ Cholesterol + Sex, data = drug, family = "binomial")
  
summary(Model.1)

```

##Model 2

```{r}

Model.2 <- glm(Drug ~ Cholesterol + Sex + BP, data = drug, family = "binomial")
  
summary(Model.2)

```


##Model 3

```{r}

Model.3 <- glm(Drug ~ Cholesterol + Sex + BP + Age, data = drug, family = "binomial")
  
summary(Model.3)

```

##Now that you've fit the three models, we will now conduct two nested model tests to determine the best of the three models. 

##Nested model test 1: Model 1 vs Model 2

```{r}

# Code to conduct a nested model test between Model 1 and Model 2 here

lrtest(Model.1, Model.2)

```

##Nested model test 2: Model 2 vs Model 3

```{r}

# Code to conduct a nested model test between Model 2 and Model 3 here

lrtest(Model.2, Model.3)

```


##Next, compute the *binarized predictions* based on the model you chose.


```{r}

# Code to create and display predicted binary outcome counts - be sure the count displays in your knitted document

# prediction
drug$predict <- predict(Model.3, newdata = drug, type = "response")
drug$predict <- ifelse(drug$predict > 0.5, 1, 0)

count(drug, predict)

```

##Confusion matrix using the table() function. Be sure to label the table axes. 

```{r}

confusion.matrix <- table("Actual" = drug$Drug, "Predicted" = drug$predict)# Complete this line
confusion.matrix

```

```{r}

# Code to compute accuracy

accuracy <- sum(diag(confusion.matrix))/sum(confusion.matrix)# Complete this line

accuracy

# Code to compute precision

precision <- confusion.matrix[2,2]/sum(confusion.matrix[,2])# Complete this line

precision

# Code to compute recall

recall <- confusion.matrix[2,2]/sum(confusion.matrix[2,])# Complete this line

recall

# Code to compute F1 score

F1score <- 2*((precision*recall)/(precision+recall))# Complete this line

F1score

```








## Check for multicollinearity between variables?

```{r}

````

## Splitting the data in train and test

```{r}
#Splitting the data using a function from dplyr package
index <- createDataPartition(drug$Drug, p = .70, list = FALSE)
train <- drug[index,]
test <- drug[-index,]

````

## Train the model

```{r}
# Training the multinomial model
multinom_model <- multinom(Drug ~ ., data = train)

# Checking the model
summary(multinom_model)
````

## Step

```{r}
exp(coef(multinom_model))
# Predicting the values for train dataset
train$DrugPredicted <- predict(multinom_model, newdata = train, "class")
# Building classification table
tab <- table(train$Drug, train$DrugPredicted)
# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(tab))/sum(tab))*100,2)
````

## Step

```{r}
# Predicting the class for test dataset
test$DrugPredicted <- predict(multinom_model, newdata = test, "class")
# Building classification table
tab <- table(test$Drug, test$DrugPredicted)
tab
````








