---
title: "Code_Handout"
author: "Jackie Vogel, Sarah Millard, Ignacio Luque, Leo Walker"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# installing caret for train/test split
library(caret)
library(nnet)
```

## Multinomial Logistic Regression

This R Markdown demonstrates how to use Multinomial Logistic Regression on a small dataset that is predicting which drug to presecribe to a patient.


```{r}
pred_maint <-read.csv("../data/predictive_maintenance.csv", header=TRUE, sep= ",")
str(pred_maint)
```

## Change the chr types to factors, limit to target = 1 so we only have occurances of failures, and then remove unnecessary variables (UID and Product_ID)


```{r}
pred_maint$Type <- as.factor(pred_maint$Type)
pred_maint$Failure_Type <- as.factor(pred_maint$Failure_Type)

pred_failure <- pred_maint[pred_maint$Target ==1,]
pred_failure <- pred_failure[pred_failure$Failure_Type != "Random Failures", ]
pred_failure <- pred_failure[,!names(pred_failure) %in% c("UDI", "Product_ID", "Target")]
```

## Check for multicollinearity between variables or other exploratory data analysis/descriptive analysis?

```{r}

````

## Splitting the data in train and test

```{r}
#Splitting the data using a function from dplyr package
index <- createDataPartition(pred_failure$Failure_Type, p = .70, list = FALSE)
train <- pred_failure[index,]
test <- pred_failure[-index,]

````

## Train the model

```{r}
# Training the multinomial model
multinom_model <- multinom(Failure_Type ~ ., data = train)

# Checking the model
summary(multinom_model)
````
## Get model coefficients

```{r}
exp(coef(multinom_model))
````

## Predicting Failure Type on Train Dataset

```{r}
# Predicting the values for train dataset
train$Failure_Predicted <- predict(multinom_model, newdata = train, "class")
# Building classification table
tab <- table(train$Failure_Type, train$Failure_Predicted)

confusionMatrix(train$Failure_Predicted, train$Failure_Type)

````

## Predicting Failure Type on Test Dataset

```{r}
# Predicting the class for test dataset
test$Failure_Predicted <- predict(multinom_model, newdata = test, "class")
# Building classification table
tab <- table(test$Failure_Type, test$Failure_Predicted)
tab
confusionMatrix(test$Failure_Predicted, test$Failure_Type)
````








